---
output:
  pdf_document: default
  html_document: default
---
\newpage

# Threshold Filtering

Threshold filters, following the guidance of Florida Department of Environmental Protection's (*FDEP*) Division of Environmental Assessment and Restoration (*DEAR*) are used to exclude specific results values from the SEACAR Analysis. Based on the threshold filters, Quality Assurance / Quality Control (*QAQC*) Flags are inserted into the *SEACAR_QAQCFlagCode* and *SEACAR_QAQC_Description* columns of the export data. The *Include* column indicates whether the *QAQC* Flag will also indicate that data are excluded from analysis. No data are excluded from the data export, but the analysis scripts can use the *Include* column to exclude data (1 to include, 0 to exclude).

```{r, results='asis', echo=FALSE, warning=FALSE}
library(kableExtra)
cont_threshold_values_df <- data.frame(
  param_name = c(
    "Dissolved Oxygen","Dissolved Oxygen","Dissolved Oxygen",
    "Salinity","Salinity","Salinity",
    "Water Temperature","Water Temperature","Water Temperature",
    "pH","pH","pH",
    "Dissolved Oxygen Saturation","Dissolved Oxygen Saturation","Dissolved Oxygen Saturation",
    "Specific Conductivity", "Specific Conductivity",
    "Turbidity","Turbidity","Turbidity"
  ),
  units = c(
    "mg/L","mg/L","mg/L",
    "ppt","ppt","ppt",
    "Degrees C","Degrees C","Degrees C",
    "pH","pH","pH",
    "%","%","%",
    "mS/cm","mS/cm",
    "NTU","NTU","NTU"
  ),
  low_thres = c(0,0,0,0,0,0,-5,-5,-5,2,2,2,0,0,0,0,0,0,0,0),
  high_thres = c(50,50,50,70,70,70,45,45,45,14,14,14,500,500,500,100,200,4000,1000,4000),
  sensor_type = c(
    "YSI EXOs","Analysis Only - 2022-04-04","6600 Series","6600 Series","YSI EXOs","Analysis Only - 2022-04-04","YSI EXOs",
    "Analysis Only - 2022-04-04","6600 Series","Analysis Only - 2022-04-04","6600 Series","YSI EXOs","YSI EXOs","6600 Series",
    "Analysis Only - 2022-04-04","6600 Series","YSI EXOs","YSI EXOs","6600 Series","Analysis Only - 2022-04-04"
  )
)

kable(cont_threshold_values_df, col.names = c("*Parameter Name*", "*Units*", "*Low Threshold*", "*High Threshold*","*Sensor Type*"), 
      escape = F, caption = "Continuous Water Quality threshold values",
      format = "simple") %>%
  kable_styling(latex_options = "scale_down",
                position = "center")
cat("\n")

```

```{r, results='asis', echo=FALSE, warning=FALSE}
library(kableExtra)
disc_threshold_values_df <- data.frame(
  param_name = c(
    "Dissolved Oxygen","Salinity","Water Temperature","pH","Dissolved Oxygen Saturation",
    "Specific Conductivity","Turbidity","Total Suspended Solids (TSS)", "Chlorophyll a uncorrected for pheophytin",
    "Chlorophyll a corrected for pheophytin", "Secchi Depth","Light Extinction Coefficient","Colored dissolved organic matter, CDOM",
    "Fluorescent dissolved organic matter, FDOM", "Total Nitrogen", "Total Kjeldahl Nitrogen TKN", "NO2+3 Filtered", "NH4 Filtered",
    "Total Phosphorus","PO4 Filtered","Ammonia- Un-ionized (NH3)", "Nitrate (N)", "Nitrite (N)", "Nitrogen, organic"
  ),
  units = c("mg/L","ppt","Degrees C","","%","mS/cm","NTU","mg/L","ug/L","ug/L","m","m^1","PCU","QSE",
            "mg/L","mg/L","mg/L","mg/L","mg/L","mg/L","mg/L","mg/L","mg/L","mg/L"
  ),
  low_thres = c("0.000001",0,3,2,"0.000001","0.005000001",0,0,0,0,"0.000001",0,0,0,0,0,0,0,0,0,0,0,0,0),
  high_thres = c(22,70,40,13,310,100,"-","-","-","-",50,"-","-","-","-","-","-","-","-","-","-","-","-","-")
)

kable(disc_threshold_values_df, col.names = c("*Parameter Name*", "*Units*", "*Low Threshold*", "*High Threshold*"), 
      escape = F, caption = "Discrete Water Quality threshold values",
      format = "simple") %>%
  kable_styling(latex_options = "scale_down",
                position = "center")
cat("\n")

```

```{r, results='asis', echo=FALSE, warning=FALSE}
threshold_df <- data.frame(desc = c("Exceeds Maximum threshold. Not verified in raw data", 
                          "Exceeds Maximum threshold. Verified in raw data",
                          "Below Minimum threshold. Not verified in raw data",
                          "Below Minimum threshold. Verified in raw data",
                          "Within threshold tolerance",
                          "No defined thresholds for this parameter"),
                 inc = c("No", "No", "No", "No", "Yes", "Yes"),
                 flag = c("2Q", "3Q", "4Q", "5Q", "6Q", "7Q"))
kable(threshold_df, col.names = c("*SEACAR QAQC Description*", "*Include*", "*SEACAR QAQCFlagCode*"),
      escape = F, caption = "Quality Assurance Flags inserted based on threshold checks listed in Table 1 & 2",
      format = "simple") %>%
  kable_styling(latex_options = "scale_down",
                position = "center")
```

# Value Qualifiers

Value qualifier codes included within the data are used to exclude certain results from the analysis. The data are retained in the data export files, but the analysis uses the *Include* column to filter the results.

**STORET and WIN value qualifier codes**

Value qualifier codes from *STORET* and *WIN* data are examined with the database and used to populate the *Include* column in data exports.

```{r, results='asis', echo=FALSE, warning=FALSE}
vq_df <- data.frame(qual_soruce = c("STORET-WIN", "STORET-WIN", "STORET-WIN", "STORET-WIN"),
                    vq = c("H","J","V","Y"),
                    inc = c("No", "No", "No", "No"), 
                    mdl = c(0, 0, 0, 0),
                    desc = c("Value based on field kit determination; results may not be accurate",
                             "Estimated value",
                             "Analyte was detected at or above method detection limit",
                             "Lab analysis from an improperly preserved sample; \ndata may be inaccurate"))

vq_df$desc = linebreak(vq_df$desc)

kable(vq_df, col.names = c("*Qualifier Source*", "*Value Qualifier*", "*Include*", "*MDL*", "*Description*"),
      escape = F, 
      caption = "Value Qualifier codes excluded from analysis",
      format = "simple",
      booktabs = T) %>%
  kable_styling(latex_options = "scale_down")
```


**Discrete Water Quality Value Qualifiers**  

The following value qualifiers are highlighted in the Discrete Water Quality section of this report. An exception is made for **Program 476** - *Charlotte Harbor Estuaries Volunteer Water Quality Monitoring Network* and data flagged with Value Qualifier **H** are included for this program only.


```{r VQ Description table long, results='asis', echo=FALSE, warning=FALSE}

# Add WQ Discrete Full-length Value Qualifier Descriptions
# add description for each VQ shown
vq <- list("N_H","N_I","N_Q","N_S","N_U")
vq_desc <- list("**H** - Value based on field kit determiniation; results may not be accurate. 
            This code shall be used if a field screening test (e.g., field gas chromatograph data, 
            immunoassay, or vendor-supplied field kit) was used to generate the value and the field 
            kit or method has not been recognized by the Department as equivalent to laboratory methods.",
                
                "**I** - The reported value is greater than or equal to the laboratory method detection 
            limit but less than the laboratory practical quantitation limit.",
                
                "**Q** - Sample held beyond the accepted holding time. This code shall be used if the value is derived 
            from a sample that was prepared or analyzed after the approved holding time restrictions for sample 
            preparation or analysis.",
                
                "**S** - Secchi disk visible to bottom of waterbody. The value reported is the depth of the waterbody 
            at the location of the Secchi disk measurement.",
                
                "**U** - Indicates that the compound was analyzed for but not detected. This symbol shall be used to indicate 
            that the specified component was not detected. The value associated with the
            qualifier shall be the laboratory method detection limit. Unless requested by the client, 
            less than the method detection limit values shall not be reported ")
vq_desc_short <- list("**H** - Value based on field kit determination",
                      "**I** - Reported value is greater than or equal to lab method detection limit, but less than quantitation limit",
                      "**Q** - Sample held beyond the accepted holding time",
                      "**S** - Secchi disk visible to bottom of waterbody",
                      "**U** - Compound was analyzed for but not detected")

vq_list <- setNames(as.list(vq_desc), vq)
vq_list_short <- setNames(as.list(vq_desc_short), vq)

cat("  \n  \n  \n")
cat("  \n")
for (vq in names(vq_list)) {
  cat(unlist(vq_list[vq]), sep = '\n')
  cat("\n")
}
cat("  \n")

```


**Systemwide Monitoring Program (SWMP) value qualifier codes**

Value qualifier codes from the *SWMP* continuous program are examined with the database and used to populate the *Include* column in data exports. *SWMP* Qualifier Codes are indicated by *QualifierSource=SWMP*.

```{r, results='asis', echo=FALSE, warning=FALSE}
swmp <- data.frame(qs = c("SWMP", "SWMP", "SWMP", "SWMP", "SWMP", "SWMP", "SWMP", "SWMP", "SWMP", "SWMP", "SWMP"),
                   vq = c(-1, -2, -3, -4, -5, 0, 1, 2, 3, 4, 5),
                   inc = c("Yes", "No", "No", "No", "No", "Yes", "No", "Yes", "Yes", "Yes", "Yes"),
                   desc = c("Optional parameter not collected","Missing data","Data rejected due to QA/QC","Outside low sensor range",
                            "Outside high sensor range","Passed initial QA/QC checks","Suspect data","Reserved for future use",
                            "Calculated data: non-vented depth/level sensor\ncorrection for changes in barometric pressure",
                            "Historical: Pre-auto QA/QC","Corrected data"))
swmp$desc <- linebreak(swmp$desc)

swmp %>% kable(col.names = c("*Qualifier Source*", "*Value Qualifier*", "*Include*", "*Description*"),
      escape = F, 
      caption = "SWMP Value Qualifier codes",
      format = "simple") %>%
  kable_styling(latex_options = c("scale_down","HOLD_position"),
                position = "center") %>% 
  row_spec(1:11,  bold = F, extra_css = 'vertical-align: top !important;')
cat("  \n")
```